Labor Markets and Technological Change: Evidence from Electronic Health Records

### 2021-10-26 Presentations/ThirdYearPaperpres.tex
Updates from Overleaf

\begin{frame}{What is an EHR?}
\begin{itemize}
    \item Computer-program technology that hospitals purchase
    \vspace{3mm}
    \item What EHR is used for depends on level purchased:
    \vspace{2mm}
    \begin{itemize}
        \item All include electronic patient records
        \vspace{2mm}
        \item More complex systems allow for decision making assistance
        \vspace{2mm}
        \item More recently, patient interfaces are being developed
    \end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}[fragile]{What does this mean for physicians?}
In the short run:
\begin{itemize}
    \item Physicians spend more time entering information into a computer and learning a new software
    \item Physicians have experienced significant burn-out from using and learning EHRs
    \item Loss of autonomy
\end{itemize}

In the long run:
\begin{itemize}
    \item Could make daily tasks easier or harder
    \item Progressing loss of autonomy as technology progresses
\end{itemize}

@@

Did EHR implementation in hospitals affect physician labor outcomes?
\begin{itemize}
    \item Intensive Margin
    \item Extensive Margin
    \item Hospital $\rightarrow$ Office?
\end{itemize}

@@

\begin{frame}{Health Information Technology}

\begin{frame}{Physician Labor}

\begin{frame}{Technology and Labor Relationship}

@@

\begin{frame}{Physician's Incentives}
\begin{itemize}
    \item Physician has a choice set of where to work, including multiple hospitals and office-based locations
    \item All physicians I consider work closely with at least one hospital at some point in time
    \item Some proportion of hospitals implement a new technology (an EHR)
    \item How to think about the effect on labor:
    \begin{enumerate}
        \item EHR makes seeing patients easier (or more difficult), thus physician can see more (or less) patients
        \item If the cost of sing an EHR is high enough, the physician can switch to working at an office-based location where they have more control
        \item A senior physician may have incentive to retire early instead of taking on the cost of learning a new technology
    \end{enumerate}
\end{itemize}
\end{frame}

@@

     Total Patients: \underline{CMS Shared Patient Data} (2009-2015)
    \begin{itemize}
        \item Sum of shared patients with all entities that have an NPI
    \end{itemize}

     Indicator for whether they are working
    \begin{itemize}
        \item =1 if Total Patients $>0$
    \end{itemize}

     Indicator for working in office-based location
    \begin{itemize}
        \item Comes from SK$\&$A database of office-based physicians
    \end{itemize}

@@

\begin{frame}{Measures of EHR Use}

Physician-level Indicator that captures exposure to \textit{any} EHR
\begin{itemize}
    \item With information on shared patients between physicians and hospitals, I limit to physicians who work in at least one hospital at some point in the time frame
    \item Hospital EHR use comes from AHA survey
    \item Similarly, I can capture hospitals that use EHRs for decision making vs. documentation using the AHA Survey IT Supplement 
\end{itemize}

@@

\begin{frame}{Summary Stats}
\centering
    \includegraphics[scale=.7]{Objects/sumstats.pdf}
\end{frame}

@@

\begin{frame}{Analysis}

@@

\begin{frame}{Full Sample Event Study}
\centering
\includegraphics[scale=.7]{Objects/event_hosp_fullsample.pdf}

>>>>

\begin{frame}{What is an EHR?}
\begin{itemize}
    \item Technology purchased and implemented by hospitals
\end{itemize}

\begin{itemize}
    \item Used for:
    \vspace{3mm}
    \begin{itemize}
        \item Electronic patient records
        \vspace{3mm}
        \item More complex (and expensive) EHRs allow for decision making assistance
        \vspace{3mm}
        \item Continuing to evolve over time
    \end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}[fragile]{What does this mean for physicians?}
Immediate Effects:
\begin{itemize}
    \item Large startup costs (time, productivity)
    \item Possible loss of autonomy
\end{itemize}

Long Run Effects:
\begin{itemize}
    \item Productivity could increase or decrease
    \item Progressing loss of autonomy as technology progresses
\end{itemize}

@@

Did EHR implementation in hospitals affect physician labor outcomes?
\begin{itemize}
    \item Intensive Margin / Productivity
    \item Extensive Margin
    \item Hospital $\rightarrow$ Office?
\end{itemize}
\end{frame}

@@

\begin{frame}{Literature}

@@

\begin{frame}{What I Want to Capture}
\begin{itemize}
    \item Initial Implementation
    \begin{itemize}
        \item Productivity gains/losses
        \item Immediate changes to labor market decisions: retirement or place of work
    \end{itemize}
    \item Loss of autonomy and cost of learning/using new technology
    \begin{itemize}
        \item Could cause physicians to move to settings where they have more control
    \end{itemize}
\end{itemize}

\begin{frame}{About the Data}
    Main measurements come from \underline{CMS Shared Patient Data} (2009-2015)
    \begin{itemize}
        \item Lists 2 NPIs and the number of billings those two entities shared in the same day
    \end{itemize}

    How I use it:
    \begin{itemize}
        \item Limit the NPIs to physician-hospital pairs, where physicians are PCPs, internists, hospitalists
        \item Limit the billings to those which occur in the same day
        \item Only keep pairs with sufficient number of billings together to suggest a close relationship between physician and hospital
        \item Aggregate to the physician level to avoid spillover effects
    \end{itemize}
\end{frame}

@@

     Total hospital patients: 
    \begin{itemize}
        \item Sum of shared patients with all hospitals
    \end{itemize}

     \vspace{4mm}

     Indicator for switching work setting:
    \begin{itemize}
        \item Positive billings shared with all entities but shared entities with hospitals becomes 0
        \item \color{gray}{SK$\&$A database of office-based physicians}
    \end{itemize}

    \vspace{4mm}

    \textcolor{gray}
     {Indicator for retirement/ leaving the labor force}

@@

\begin{frame}{Measure of EHR Use}

Physician-level treatment variable that captures exposure to \textit{any} EHR
\begin{itemize}
    \item Individual hospital EHR use comes from AHA survey
    \item Treatment turns on the first time any of the shared hospitals implements an EHR
\end{itemize}
\end{frame}

@@

\begin{frame}{Summary Statistics}
\centering
    \includegraphics[scale=.8]{Objects/sumstats.pdf}
\end{frame}

@@

\begin{frame}{Event Studies}
\begin{equation*}
    Y_{it}=\alpha_t+\delta_i
\end{equation*}

@@

\begin{frame}{Intensive Margin}
\centering
\includegraphics[scale=.45]{Objects/event_plot.pdf}
\end{frame}

\begin{frame}{Intensive Margin}
\centering
\includegraphics[scale=.45]{Objects/event_hosp_plot.pdf}
\end{frame}

\begin{frame}{Hospital to Office?}
    \centering
    \includegraphics[scale=.5]{Objects/event_plot_office_to_hosp.pdf}
\end{frame}

\begin{frame}{Hospital to Office?}
    \centering
    \includegraphics[scale=.5]{Objects/event_hosp_plot_hosp_to_office.pdf}
\end{frame}
